---
title: "Group Project"
output: html_document
date: "2025-11-19"
---

## Setup Data

Load data

```{r}
library(dplyr)
library(ggplot2)

data = read.csv("song_data.csv")
head(data)
```

Drop the `song_name` column as this is just for identifying the song.
```{r}
to_drop = c("song_name")
data = data[, !(names(data) %in% to_drop)]
```

Get the column names
```{r}
column_names = colnames(data)
print(column_names)
```

From the description of the data, we know that `key`, `audio_mode` and `time_signature` are categorial data. So we can split the column names into categorial and continuous data
```{r}
categorial_col_names = c("key", "audio_mode", "time_signature")
continuous_col_names = column_names[!column_names %in% categorial_col_names]

continuous_data = data[, (names(data) %in% continuous_col_names)]
categorial_data = data[, (names(data) %in% categorial_col_names)]

cat("Categorial: ", categorial_col_names, "\n")
cat("Continuous: ", continuous_col_names, "\n")
```

## Data Cleaning

Check if there are duplicate songs in our dataset
```{r}
length(unique(data)) == nrow(data)
```

Since the number of unique songs is not the same as the number of rows in data, we can see that there are duplicate data in our dataset.

Drop the exact duplicate data.
```{r}
data = unique(data)
```

Check for missing data.
```{r}
which(is.na(data))
```

Since integer(0) is being returned, this means that there is no missing data to clean.

Check values and frequency of the categorial data
```{r}
for (name in categorial_col_names) {
  frequency_table = table(data[[name]])
  print(name)
  print(frequency_table)
  print(prop.table(frequency_table))
  cat("\n")
}
```

We can see that there is no inconsistent data for the categorial data.

Apply factors to the categorial data
```{r}
for (name in categorial_col_names) {
  data[[name]] = factor(data[[name]])
}
```

## Data Exploration

Display the mean, median, and standard deviation of each of the continuous data
```{r}
for (name in continuous_col_names) {
  curr_col = data[[name]]
  
  curr_col_mean = mean(curr_col)
  curr_col_median = median(curr_col)
  curr_col_sd = sd(curr_col)

  cat(name, "\n")
  cat("Mean: ", curr_col_mean, "\n")
  cat("Median: ", curr_col_median, "\n")
  cat("Standard Deviation", curr_col_sd, "\n");
  cat("\n")
}
```

Plot a density of each continuous column
```{r}
for (name in continuous_col_names) {
  print(ggplot(data, aes(x=data[[name]])) + geom_histogram(aes(y=..density..)) + geom_density(linewidth = 1) + labs(title=paste("Histogram of", name)))
}
```

Plot a bar graph for each categorial data
```{r}
for (name in categorial_col_names) {
  print(ggplot(data, aes(x=data[[name]])) + geom_bar() + labs(title=paste("Count of", name)))
}
```

Display variance-covariance matrix
```{r}
varcov_matrix = cov(continuous_data)
print(varcov_matrix)
```

Display correlation matrix
```{r}
cor_matrix = cor(continuous_data)
print(cor_matrix)
```

Standarize data
```{r}
standardized_data = as.data.frame(scale(data[, continuous_col_names]))
```

Split into training and testing data
```{r}
set.seed(123456789)
data_size = nrow(standardized_data)
training_data_size = floor(data_size * 0.8)
training_index = sample(seq_len(data_size), size=training_data_size)

training_data = standardized_data[training_index,]
testing_data = standardized_data[-training_index,]
```
















